# 경영통계분석 HW1, R code {.unnumbered}

20249132 김형환

## 2. Discrete data

```{r}
#| output: false
library(tidyverse)
birds_caribbean <- tibble("dates"=c(0.00,0.00,0.04,0.21,0.29,
0.54,0.63,0.88,0.96,1.25,1.67,1.75,1.84,1.96,2.01,
2.51,2.72,3.30,3.51,4.05,4.85,6.94,8.73,10.57,11.11,
12.45,14.00,17.30,17.92,18.05,18.43,22.48,22.48,23.48,26.32,
26.45,28.87))
```

**a.  Plot the data in a histogram and describe the shape of the frequency distribution.**

```{r, fig.width=10,fig.height=4}
ggplot(data=birds_caribbean, aes(x=dates))+
  geom_histogram(binwidth=5, boundary=0, color="black",fill="darkred")+
  scale_x_continuous(breaks = seq(0,30,5)) +
  theme_bw()
```

**c.  Calculate the mean and median. Was your intuition in part (b) correct?**

```{r}
birds_caribbean %>% 
  summarise(mean=mean(dates), mid=median(dates))
```

**d.  Calculate the first and third quantiles and the inter quartile range.**

```{r}
birds_caribbean %>% reframe(quantile=quantile(dates))
```

**e.  Draw a box plot for these data**

```{r, fig.width=10,fig.height=2}
ggplot(birds_caribbean, aes(x=dates))+
  geom_boxplot()+
  scale_y_continuous(limits=c(-0.5,0.5))+
  theme_bw()
```

## 4. Sample Statistics

```{r}
handheight <- tibble("person"=c("A","B","C","D","E"),
                     "hand"=c(17,15,19,17,21),
                     "height"=c(150,154,169,172,175))
handheight
```

**a. Calculate the sample variances for hand and height, respectively.**

```{r}
sample_var <- handheight %>% 
  mutate(hand_tmp=hand-mean(handheight$hand),
         height_tmp=height-mean(handheight$height)) %>% 
  mutate(hand_tmp=hand_tmp^2, height_tmp=height_tmp^2) %>% 
  summarise(s_var_hand=sum(hand_tmp)/4, s_var_height=sum(height_tmp)/4)
sample_var
```

**b. Calculate the sample covariance.**

```{r}
sample_cov <- handheight %>% 
  mutate(s_cov_tmp=(hand-mean(handheight$hand))*(height-mean(handheight$height))) %>% 
  summarise(s_cov=sum(s_cov_tmp)/4)
sample_cov
```

**c. Calculate the sample correlation and interpret the result.**

```{r}
sample_corr=sample_cov$s_cov/sqrt(sample_var$s_var_hand*sample_var$s_var_height)
sample_corr
```

::: {.callout-tip title="R 내장함수 활용 가능"}
```{r}
var(handheight$hand)
var(handheight$height)
cov(handheight$hand,handheight$height)
cor(handheight$hand,handheight$height)
```
:::